P4IRS: An Intermediate Representation and Compiler for Parallel and Performance-Portable Particle Simulations

### 2025-05-02 sage_latex_template/manuscript.tex
minor
Signed-off-by: Rafael Ravedutti <rafael.r.ravedutti@fau.de>
minor hl
Signed-off-by: Rafael Ravedutti <rafael.r.ravedutti@fau.de>
Add url for Jinja templates
Signed-off-by: Rafael Ravedutti <rafael.r.ravedutti@fau.de>

MESA-PD (\cite{mesapd3}) is a general particle dynamics framework that is built-in the waLBerla framework for the simulation of particles and fluids.
Its design principle of separation of code and data allows to introduce arbitrary interaction models with ease, which makes it also useful for \ac{MD}.
By using its code generation approach through Jinja templates, it can be adapted very effectively to any simulation scenario.
As successor of the ``physics engine'' (pe) rigid particle dynamics framework it inherits its scalability and advanced load balancing functionalities.
One of the limitations of MESA-PD is that it does not support GPU code generation, and since its code generation approach is based on Jinja templates, the effort for implementing new functionalities may not be lower compared to other simulation packages without code generation techniques (for instance, most of its force kernels are not generated but rather implemented in C++).

@@

\autoref{lst:pairsruntime} shows an example for the routine to write data into VTK files, note that the first parameter is a \emph{PairsRuntime} object, and can be used to check whether properties exist, check if they are synced with the host device and (in case not) copy the data from the host.
The meta-data object is feed at the start of the simulation with the code generated by P4IRS, and it is therefore dependent on each simulation.
It is important to enhance that accessing simulation data through the \emph{PairsRuntime} instance is not as efficient as accessing them directly in the generated code since the way of accessing the data structures (such as the property layouts) must be evaluated during runtime.
Therefore, the usage of such instance must be avoided for performance-intensive routines.

>>>>

MESA-PD (\cite{mesapd3}) is a general particle dynamics framework that is built-in the waLBerla framework for the simulation of particles and fluids.
Its design principle of separation of code and data allows to introduce arbitrary interaction models with ease, which makes it also useful for \ac{MD}.
By using its code generation approach through Jinja templates\footnote{\url{https://palletsprojects.com/projects/jinja/}}, it can be adapted very effectively to any simulation scenario.
As successor of the ``physics engine'' (pe) rigid particle dynamics framework it inherits its scalability and advanced load balancing functionalities.
One of the limitations of MESA-PD is that it does not support GPU code generation, and since its code generation approach is based on Jinja templates, the effort for implementing new functionalities may not be lower compared to other simulation packages without code generation techniques (for instance, most of its force kernels are not generated but rather implemented in C++).

@@

\autoref{lst:pairsruntime} shows an example for the routine to write data into VTK files, note that the first parameter is a \emph{PairsRuntime} object, and can be used to check whether properties exist, check if they are synced with the host device and (in case not) copy the data from the host.
The meta-data object is \RMchange{fed} at the start of the simulation with the code generated by P4IRS, and it is therefore dependent on each simulation.
It is important to enhance that accessing simulation data through the \emph{PairsRuntime} instance is not as efficient as accessing them directly in the generated code since the way of accessing the data structures (such as the property layouts) must be evaluated during runtime.
Therefore, the usage of such instance must be avoided for performance-intensive routines.
